Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-03-24 Publications/2015-NeuroImage/02-abstract.tex
Updates after Yasser and Dominique

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
\emph{Regseg} showed a consistent error within the 95\% confidence interval of
  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.

>>>>

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
We demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
  respect to a ground truth, obtaining a 95\% confidence interval of 0.56 - 0.66 (mm)
  for the Hausdorff distance of mesh vertices in images of 1.25mm isotropic resolution.

### 2015-03-24 Publications/2015-NeuroImage/03-introduction.tex
Updates after Yasser and Dominique

% -*- root: 00-main.tex -*-
\section{Introduction}%
\label{sec:introduction}
Image registration is the process to find a spatial mapping that aligns the information available
  in two different coordinate systems.
In their comprehensive survey, \cite{sotiras_deformable_2013} illustrated the wide variety
  of existing methodologies, and classified them by their three principal building blocks:
  the matching criteria to evaluate the proximity of the solution,
  the deformation model that may implement theoretical properties of the mapping
  (i.e. linearity, elasticity, viscosity, etc.), and the optimization method.
The matching criteria can be defined as a function on the intensities of voxels of the
  two images, the alignment of spatial features derived from the corresponding objects in
  the spaces to be registered, or use both intensity and spatial informations.
In \citep{sotiras_deformable_2013} these methods are referred to as \emph{iconic}, \emph{geometric},
  and \emph{hybrid}, respectively.
When the matching criteria is defined over an optimizable partition of one or both reference an
  moving spaces, we face a hybrid method which simultaneously segments and registers images.
An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
  a sequential deformation of active contours for object tracking in series of 2D images.
Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
  the registration problem with an affine transformation supporting the mapping.
\cite{vemuri_joint_2003} proposed an atlas-based registration framework using level sets and only
  one \gls*{pde} for first time.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2003}
  to nonlinear registration implementing a free deformation field.
\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration and segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
\cite{greve_accurate_2009} presented a widely used registration method called \emph{bbregister},
  included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
Their framework performs robust registration of brain surfaces into the intensity information
  of a target \gls*{mri}, using an affine transformation and active contours.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method using level sets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves topology even with very large deformations.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  level-set function for the registration of several active contours, in the application
  of atlas-based segmentation.
Alternatively to the historical use of active contours, some Bayesian approaches
  have been proposed as well \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.

We propose \emph{regseg}, an active-contours driven registration method.
Then, we demonstrate its aptness on the correction of the nonlinear distortions typically
  presented by \gls*{dmri} data of the brain.
This problem is briefly described in \autoref{sec:distortions}.
The general overview of our method, along with the contributions with respect 
  to the current state of art are summarized in \autoref{sec:contributions}.
In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show its duality
  with respect to the active-contours route.
Following sections
  introduce the specific features of \emph{regseg} (\autoref{sec:numerical_implementation}),
  state the evaluation experiments (\autoref{sec:experiments_evaluation}),
  and describe the synthetic and real data involved in this study (\autoref{sec:datasets}).
In \autoref{sec:results} we describe our findings to support the usefulness of \emph{regseg}
  in the described application, compared to a well-established tool.
The interpretation of the results and prospects is addressed in \autoref{sec:discussion}.

>>>>

% -*- root: 00-main.tex -*-
\section{Introduction}%
\label{sec:introduction}
Image registration and segmentation are crucial in current pipelines for
  analysis of whole-brain connectivity using \gls*{dmri}, such as the extraction
  of brain connectivity \citep{craddock_imaging_2013}.

An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
  a sequential deformation of active contours for object tracking in series of 2D images.
Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
  the registration problem with an affine transformation supporting the mapping.
\cite{vemuri_joint_2003} proposed an atlas-based registration framework using level sets and only
  one \gls*{pde} for first time.
\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2003}
  to nonlinear registration implementing a free deformation field.
\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
  registration and segmentation, through the propagation of the deformation field from
  the contours to the whole image definition.
\cite{greve_accurate_2009} presented a widely used registration method called \emph{bbregister},
  included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
Their framework performs robust registration of brain surfaces into the intensity information
  of a target \gls*{mri}, using an affine transformation and active contours.
Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method using level sets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves topology even with very large deformations.
Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
  level-set function for the registration of several active contours, in the application
  of atlas-based segmentation.
Alternatively to the historical use of active contours, some Bayesian approaches
  have been proposed as well \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.

We propose \emph{regseg} as a solution to joint segmentation and registration 
  problems with a shape priors approach.
An immediate application of \emph{regseg} is the correction of susceptibility-derived
  distortions typically present in \gls*{dmri} data and its segmentation
  after correction.
Both correction and segmentation are required steps in current investigations of whole-brain
  tractography, such as the connectome extraction \citep{daducci_connectome_2012}.
Indeed, the joint segmentation and correction can exploit the anatomy extracted from
  the \gls*{t1} as strong shape-priors.

### 2015-03-24 Publications/2015-NeuroImage/04-methods.tex
editions after @daducci review
Updates after Yasser and Dominique

\subsection{Susceptibility-derived distortions}
\label{sec:distortions}
\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
In order to quickly probe the diffusion process within the brain in many different orientations,
  these sequences minimize the bandwidth of the \gls*{pe} direction.
This low bandwidth originates the artifactual displacement of locations presenting small deviations
  from the main magnetic field $B_0$.
One important source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
  susceptibility across the imaged object.
Tissue/air boundaries show large steps of susceptibility and this artifact is
  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
  bone of both hemispheres.
Distortion can be rendered as a registration problem between the undistorted $R$ (reference), 
  and distorted $M$ (moving) spaces:

  \begin{align}
  U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}
%
  where $\vec{r}$ denotes a position in $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ the dimensionality of images.
Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
  to the reference domain.

The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
  \citep{jezzard_correction_1995}:

  \begin{equation}
  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
  \label{eq:fieldmap}
  \end{equation}
%
where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
  vector along \gls*{pe}.
Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
The first solution consisted on estimating $\Delta B_0$ with field mapping
  techniques \citep{andersson_modeling_2001}.
A second family of methods require an extra \gls*{epi} acquisition, but switching
  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
  image\footnote{The term \emph{b0} refers to one or more images within the
  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
   \emph{b}-value.} of the \gls*{dmri} to an undistorted \gls*{t2} image of
   the same subject.
In the following, we will refer to this method as \gls*{t2b}.
Further developments \citep{irfanoglu_susceptibility_2011} extended the \gls*{t2b} 
  approach to the 3D problem.
Therefore, a typical protocol for the extraction of structural connectivity
  comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
  method of choice \citep{daducci_connectome_2012}.

\subsection{Our method and contributions}
\label{sec:contributions}
%
\emph{Regseg} can be viewed as a derivation of \emph{bbregister} \citep{greve_accurate_2009}.
Both methods exploit the anatomically correct and precise surfaces that are readily
  extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}, mapping them
  to a target \gls*{dmri} image using active contours.
We implemented active contours without edges \citep{chan_active_2001}, that allow
  \emph{regseg} to segment homogeneous regions within the brain, while \emph{bbregister}
  use active contours with edges to look for intensity boundaries in the target image.
Another identifying difference is the deformation model:
  \emph{bbregister} includes an affine transformation and overcomes the distortion
  problem by excluding typically affected regions from computation.
Conversely, \emph{regseg} implements a free-deformation field supported with B-Spline
  functions.
Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
  in the use of active contours without edges for registration, however we avoided
  the use of level sets with the application of shape gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
\cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
  proposed a nonlinear elasticity smoother.
In contrast, \emph{regseg} implements the anisotropic regularization proposed by
  \citep{nagel_investigation_1986}.
We also present an ellaborated evaluation instrumentation to create realistic distortions
  and test algorithms.
Integrating \emph{regseg} and an in-house implementation of the \gls*{t2b} correction
  in the workflow, we compared the performance of both solutions.

@@

  \begin{equation}
  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}
%
  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
  between piecewise smooth regions $\Omega_i$, such that
  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
  contour between competing regions $\Omega_l$ and $\Omega_m$.
The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.

First, we assume independence between pixels, and thus break down the
  global data likelihood into a product of pixel-wise conditional probabilities:

  \begin{equation}
  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
%
  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.

@@

  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= C + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}
%
  this expression is dual to the energy functional corresponding
  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
  with the corresponding anisotropic regularization term \citep{nagel_investigation_1986}.

@@

Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$:

@@

In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3 ($\beta_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
  $\psi$ by the actual kernel function, the transformation writes:

  \begin{equation}
    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
  \label{eq:transformation}
  \end{equation}
%
  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

@@

\paragraph*{Simulated phantoms}%
\label{sec:digital_phantoms}
We proved the concept on simplistic digital phantoms which we will name after their
  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
  boxes A and B).
Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated
  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
  corresponding to each phantom type, with two resolutions each
  ($1.0mm$ and $2.0mm$ isotropic).
Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
The reference surfaces were extracted at the higher resolution (1.0mm isotropic),
  using \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
  \emph{FreeSurfer}.}.

@@

To warp the simulated datasets, we generated 150 different realizations of
  the distortion $U_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
The coefficients of the B-Spline functions were randomly generated for each resolution level.
Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

>>>>

\subsection{Susceptibility-derived distortions}
\label{sec:distortions}
\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
In order to quickly probe the diffusion process within the brain in many different orientations,
  these sequences minimize the bandwidth of the \gls*{pe} direction.
This limitation originates a warping of the reconstructed image in regions affected by
  small deviations from the main magnetic field $B_0$.
One important source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
  susceptibility across the imaged object.
Tissue/air boundaries show large steps of susceptibility and this artifact is
  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
  bone of both hemispheres.

Distortion can be rendered as a registration problem where we look for an underlying deformation 
  field $U_s$, mapping from an undistorted \gls*{t1} image (the reference $R$) to the distorted
  \gls*{dmri} images (the moving space $M$), such that the structures identified in $R$ optimally
  align with the corresponding features in $M$:

  \begin{align}
  U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
  \label{eq:transform}
  \end{align}
%
  where $\vec{r}$ denotes a position in $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ the dimensionality of images.
Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
  to the reference domain.
The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
  \citep{jezzard_correction_1995}:

  \begin{equation}
  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
  \label{eq:fieldmap}
  \end{equation}
%
where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
  vector along \gls*{pe}.
Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
The first solution consisted on estimating $\Delta B_0$ with field mapping
  techniques \citep{andersson_modeling_2001}.
A second family of methods require an extra \gls*{epi} acquisition, but switching
  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
  image\footnote{The term \emph{b0} refers to one or more images within the
  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
   \emph{b}-value.} of the \gls*{dmri} to an undistorted \gls*{t2} image of
   the same subject.
In the following, we will refer to this method as \gls*{t2b}.
Further developments \citep{irfanoglu_susceptibility_2011} extended the \gls*{t2b} 
  approach to the 3D problem.

\subsection{Contributions}
\label{sec:contributions}
%
While all the correction methods require an extra acquisition, \emph{regseg} is especially
  indicated in scanning protocols without specific data for correction,
  such as historical datasets.

Joint segmentation and distortion correction of \gls*{dmri} data is also interesting to
  improve the overall consistence of complex processing workflows.
The closest antecedent of \emph{regseg} is \emph{bbregister} \citep{greve_accurate_2009}.
Both methods use active contours to align surfaces to same-subject \gls*{dmri} images.
However, \emph{bbregister} is limited to affine transformations, and thus, it is not able to
  solve the susceptibility-derived distortions, generally nonlinear.
To enable registration of uncorrected data, \emph{bbregister} excludes from the cost-function
  those regions frequently warped.
Conversely, \emph{regseg} implements a free-deformation field, supported with B-Spline
  functions to estimate deformations.
A second diverging foundation is the active-contours framework.
In \emph{regseg}, active contours are implemented without edges \citep{chan_active_2001},
  enabling the tool to segment homogeneous regions within the brain.
On the other hand, \emph{bbregister} requires that the target features present intensity
  boundaries as it implements active contours with edges.
The tissue interfaces imaged with \gls*{dmri} are affected by partial volume effects that
  blur boundaries.
In practice, registration of \gls*{t1} and \gls*{dmri} data with \emph{bbregister}
  is limited to the linear match of the pial surface of the brain.

Finally, we present an instrumentation framework to evaluate correction and
  segmentation pipelines using realistic distortions as ground truth.
With the integration of \emph{regseg} and an in-house implementation of the 
  \gls*{t2b} correction in the framework, we compare the performance
  of both solutions.

% Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
%   in the use of active contours without edges for registration, however we avoided
%   the use of level sets with the application of shape gradients
%   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
% \cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
%   proposed a nonlinear elasticity smoother.
% In contrast, \emph{regseg} implements the anisotropic regularization proposed by
%   \citep{nagel_investigation_1986}.
% We demonstrate its aptness on the correction of the nonlinear distortions typically
%   presented by \gls*{dmri} data of the brain.
% Therefore, a typical protocol for the extraction of structural connectivity
%   comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
%   method of choice \citep{daducci_connectome_2012}.

@@

  \begin{equation}
  P(U \mid R, \boldsymbol{\Gamma}) = \frac{P(R \mid U,\boldsymbol{\Gamma})\, P(U)}{P(R)},
  \label{eq:bayes_rule}
  \end{equation}
%
  where $P(R \mid U,\boldsymbol{\Gamma})$ is the data-likelihood, and
  $\boldsymbol{\Gamma} = \{\Gamma_{\text{in},\text{out}}\}$ is a set of surfaces 
  corresponding to the interfaces between pairs of competing regions $\Omega_\text{in}$
  and $\Omega_\text{out}$ drawn from the full set of regions $\{\Omega_l \mid l \in \mathbb{N}, l \leq L\}$.
As $\boldsymbol{\Gamma}$ are mapped by $U$, we simplify
  $P(U, \boldsymbol{\Gamma}) = P(U) \implies P(R \mid U,\boldsymbol{\Gamma}) = P(R \mid U)$.

The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
  \citep{bishop_pattern_2006}, and aligns $\boldsymbol{\Gamma}$ into $R$.
First, we assume independence between pixels, and thus break down the
  global data likelihood into a product of pixel-wise conditional probabilities:

  \begin{equation}
  P(R \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:bayes_aposteriori}
  \end{equation}
%
  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
  volume.
For convenience, and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
  formulation of the \emph{data term} of our registration model:

 	\begin{align}
  P( R \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the reference image $R$.

@@

  \begin{align}
  E(R \mid U) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= \const + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \notag\\
  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r},
  \label{eq:energy}
  \end{align}
%
  this expression is dual to the energy functional corresponding
  to a discrete \acrlong*{acwe} framework \citep{chan_active_2001}
  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.

@@

Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved through a set of associated basis functions $\psi_k$.
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
  of degree 3: % ($\beta_3$).

@@

%Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
%  $\psi$ by the actual kernel function, the transformation writes:
%
%  \begin{equation}
%    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
%      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
%  \label{eq:transformation}
%  \end{equation}
%%
%  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.

@@

\paragraph*{Simulated phantoms}%
\label{sec:digital_phantoms}
We proved the concept on simplistic digital phantoms which we will name after their
  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
  boxes A and B).
Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated
  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
  corresponding to each phantom type, with two resolutions each
  (1.0mm and 2.0mm isotropic).
Field of view at both resolutions was \isores{100}mm.
Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
The reference surfaces were extracted at high resolution, using 
  \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
  \emph{FreeSurfer}.}.

@@

To warp the simulated datasets, we generated 150 different realizations of
  the distortion $U_{true}$.
Smoothness is introduced using two levels of B-Spline functions, with control points evenly
  located in isotropic grids covering the full extent of the phantom.
Separation between control points was isotropic, and set to 50.5mm for the first level
  of distortion and 25.25mm for the second.
The rationale behind this choice is producing large deformations on the contours with
  the first level, and matching the properties of susceptibility-derived distortions
  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
The coefficients of the B-Spline functions were randomly generated for each resolution level.
Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
  level (20.2mm and 10.1mm respectively) as studied in \citep{rueckert_diffeomorphic_2006}.

### 2015-03-24 Publications/2015-NeuroImage/08-acknowledgements.tex
Updates after Yasser and Dominique

DZ is supported by the Swiss National Science Foundation under grant PBELP2-137727.
This study is supported by: the Spain's Ministry of Science and Innovation
  (projects TEC2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
  European Regional Development Funds; the Center for Biomedical Imaging
  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
  Leenaards and Louis Jeantet foundations.

>>>>

DZ was supported by the Swiss National Science Foundation under grants PBELP2\_137727 
  and P300P2\_147778, and NSF-DMS 1418812
This study is supported by: the Spain's Ministry of Science and Innovation
  (projects TEC2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
  European Regional Development Funds; the Center for Biomedical Imaging
  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
  Leenaards and Louis Jeantet foundations.
